---
title: "Cobb Douglas Production Function: Non Linear Optimization"
author: "Carlos Pelayo"
date: "4/24/2020"
output: html_document
---

## non-linear optimization

This Rmarkdown file demonstrates techniques for non-linear optimization, and is illustrated with data for the Cobb-Douglas production function, $y = A K^\alpha L^\beta$, K capital, L labor, and the exponents the elasticies of production by input.

### Methods demonstrated:
 1. Grid search of space of parameters  
 2. non-linear optimization software, find values of the parameters that minimize the sum of squared errors, errors are (data-predicted)  
 3. non-linear optimizatin software, maximum likelihood- find values of the parameters that maximize the likelihood of the data assuming errors are normally distributed  
 4. log-transform to a linear model  
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(scatterplot3d)
```

Read in the data

```{r}
CobbDoug = read.csv('CobbDouglasData.csv')
attach(CobbDoug)
```


```{r}
plot(Cap, Q, pch=19, col='red', xlab= 'input values', ylab = 'output',
     main='Cobb-Douglas Production function')
points(Emp, Q, pch=19, col='blue')
grid()
```

A two-dimensional plot would be nice..

```{r}
scatterplot3d(x=Cap,y=Emp,z=Q,color="steelblue",pch="$",angle=15, main="Cobb Douglas Production Function", xlab="Production",ylab="Labor Input",zlab="Capital Input",grid=TRUE)

```

Set up a non-linear loss function

```{r}
CDloss=function(parms, Q, Cap, Emp){
  
  A = parms[1]
  alpha = parms[2]
  beta = parms[3]
  
  Loss = 0
  for (i in 1:length(Q)){
    Loss = Loss + ((Q[i]-A*(Cap[i]^beta*Emp[i]^alpha))^2)
  }
  
  return (log(Loss))
}

CDmle=function(parms, Q, Cap, Emp){
  A = parms[1]
  alpha = parms[2]
  beta = parms[3]
  
  logLoss = 0
  for (i in 1:length(Q)){
    errorTerm= (Q[i]-A*(Cap[i]^beta*Emp[i]^alpha))
    logLoss = logLoss - dnorm(errorTerm, log=T)
  }
  
  return (log(logLoss))
}
```

Test the loss function

```{r}
A = 1
alpha = 0.5
beta = 0.5
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))

# better or worse than
A = 0.75
alpha = 0.2
beta = 0.5
parms=c(A, alpha, beta)
print(CDloss(parms, Q, Cap, Emp))

```
## Grid search
  not the best idea, can get close to an answer 

```{r}

for (A in seq(from=0.1, to = 3, length.out = 4))
  for(a in seq(from=0.1, to = 1,length.out = 4))
   for(b in seq(from=0.1, to = 1, length.out = 4)){
     parms = c(A, a, b)
      print(paste(A,"/" ,a,"/", b,"/", CDloss(parms, Q, Cap, Emp)))
  }

```

## non linear optimization

use function optim to find values of A, alpha, beta
find values that minimize the loss function, loss = sum of sqrared errors.  This is a 'least squares' solution

```{r}
### Optimization
parms = c(1, 0.75, 0.25)
opt = optim(parms, CDloss,Q=Q,Cap=Cap,Emp=Emp)
print(opt)
parms = c(opt$par[1], opt$par[2], opt$par[3])
print(CDloss(parms, Q, Cap,Emp))
```

## transform to logs

```{r}
logModel = glm(log(Q)~log(Cap)+log(Emp))
summary(logModel)
plot(logModel)
A = exp(logModel$coefficients[1])
alpha = logModel$coefficients[2]
beta = logModel$coefficients[3]
print(paste('A = ',A,' alpha = ', alpha,' beta =', beta))

# CD loss at these values:
parms=c(A, alpha, beta)
print(paste('CD loss: ', CDloss(parms, Q, Cap, Emp)))
```
$$log(Q)=0.54+0.25*log(k)+0.77*log(L)$$
$$Q=e^{0.54}*K^{0.23}+L^{0.77}$$

## maximum likelihood estimation

Wi 

$\phi$ is the standard normal distribution


```{r}
hist(logModel$residuals)
parms = c(2, 0.8, 0.2)
opt = optim(parms, CDmle, Q=Q, Cap=Cap, Emp=Emp)
print(opt)
parms = c(opt$par[1], opt$par[2], opt$par[3])
print(CDmle(parms, Q, Cap,Emp))

```

## Questions:
### 1. what are the estimated values of A, alpha and beta?
    In order to minimize the loss function, the values for the constants are as follow: $A$=1.3044599, $\alpha$=0.8287067, $\beta$=0.2215366. The result was acheieved by using the Maximum Likelihood estimation method.

### 2. Confidence intervals on A, alpha, beta?

  Confidence Interval for $\alpha$: 0.23 $\pm$ 1.96*0.02882

  Confidence Interval for $\beta$: 0.77 $\pm$ 1.96*0.02735

  Confidence Interval for $A$: 1.716$\pm$ $e^{0.2960}$

### 3. Is production at constant return to scale?
  If production is increasing or decreasing return to scale, which input is most  expensive? Production is at a constant return to scale. It is contstant because $\alpha$+$\beta$ is equal to 0.9976 which is almost equal to 1.

### 4. What does the values of A mean?
  $A$ is the total factor productivity, which is an efficiency parameter.

